### 2020-02-14 mpc.Rnw

\footnotesize 
\begin{tabular}{@{\extracolsep{5pt}}lcccc} 
\\[-1.8ex]\hline 
\hline 
\\[-1.8ex] & \multicolumn{1}{c}{Full Sample} & \multicolumn{1}{c}{Resident}& \multicolumn{1}{c}{NSFH}& \multicolumn{1}{c}{NSFG} \Writing to file mpc.tex
Processing code chunks with options ...

You can now run (pdf)latex on ‘mpc.tex’
}{c}{(4)}\\ 
\hline \\[-1.8ex] 
\\[-2.2ex] & \multicolumn{4}{c}{Risk of Marriage relative to Cohabitation} \\  
 \hline \\[-1.8ex]
 Unilateral Divorce &  -0.09  &  -0.07  &  -0.12  &  -0.13  \\ 
  & ( 0.04 ) & ( 0.04 ) & ( 0.04 ) & ( 0.11 ) \\  
 \hline \\[-1.8ex]
 Average Relative Risk &  -0.12  &  -0.08  &  -0.2  &  -0.14  \\ 
 \hline \\[-1.8ex]
 \\[-2.2ex] & \multicolumn{4}{c}{Risk of Separation relative to Cohabitation} \\  
 \hline \\[-1.8ex]
 Unilateral Divorce &  -0.04  &  -0.09  &  -0.05  &  -0.06  \\ 
  & ( 0.02 ) & ( 0.02 ) & ( 0.02 ) & ( 0.1 ) \\  
 \hline \\[-1.8ex]
 Average Relative Risk &  -0.06  &  -0.18  &  -0.1  &  -0.06  \\ 
 \hline \\[-1.8ex]
State Fixed effects & Yes & Yes & Yes & Yes \\ 
Year Fixed effects & Yes & Yes & Yes & Yes \\ 
Age Polynomial & Yes & Yes & Yes & Yes \\
Picewise Duration & Yes & Yes & Yes & Yes \\ 
\hline
Observations & \multicolumn{1}{c}{ 138012 } & \multicolumn{1}{c}{ 81920 } & \multicolumn{1}{c}{ 77826 } & \multicolumn{1}{c}{ 60186 } \\ 
\hline
Censored spells(\%) & \multicolumn{1}{c}{ 10.18 } & \multicolumn{1}{c}{ 10.98 } & \multicolumn{1}{c}{ 11.6 } & \multicolumn{1}{c}{ 8.38 } \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}