### 2020-02-14 inc_men.tex
only results left

{\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}              \begin{tabular}{l*{1}{c}}                          \toprule             &\multicolumn{1}{c}{(1)}       \\             \midrule             \textsc{Dep. Variable:} &  \\\textsc{Male Log Earnings} & \\ & \\
$\iota^m_1$         &        0.05\\
$\iota^m_2$         &       -0.00\\
$\iota^m_0$           &       -0.34\\
Survey Year Fixed Effects  & \checkmark   \\         State Fixed Effects  & \checkmark  \\                         \hline
Observations        &       98118\\
\(R^{2}\)           &       0.152\\
\hline