### 2020-02-14 inc_wom.tex

{\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}              \begin{tabular}{l*{1}{c}}                          \toprule             &\multicolumn{1}{c}{(1)}       \\             \midrule             \textsc{Dep. Variable:} &  \\\textsc{Female Labor Earnings} & \\ & \\
$\iota^f_1$         &        0.02\\
$\iota^f_2$         &       -0.00\\
$\iota^f_0$            &       -0.38\\
Survey Year Fixed Effects  & \checkmark   \\         State Fixed Effects  & \checkmark  \\                         \hline
Observations        &       86891\\
\(R^{2}\)           &       0.085\\
\hline