### 2020-02-14 prb_wom.tex

{\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}              \begin{tabular}{l*{1}{c}}                          \toprule             &\multicolumn{1}{c}{(1)}       \\             \midrule             \textsc{Dep. Variable:} &  \\\textsc{Female Labor Force Participation} & \\ & \\
Unilateral Divorce$*$Community Property&       -0.18\sym{***}\\
Unilateral Divorce$*$Title Based&       -0.08         \\
Unilateral Divorce$*$Equitable Distribution&       -0.06         \\
Equitable Distribution&       -0.00         \\
$\iota^f_1$         &        0.01\sym{***}\\
$\iota^f_2$         &       -0.00\sym{***}\\
$\iota^f_0$            &        1.95         \\
Survey Year Fixed Effects  & \checkmark   \\         State Fixed Effects  & \checkmark  \\                         \hline
Observations        &      127728         \\
\hline