### 2020-09-28 mpc.tex

\footnotesize 
\begin{tabular}{@{\extracolsep{5pt}}lcccc} 
\\[-1.8ex]\hline 
\hline 
\\[-1.8ex] & \multicolumn{1}{c}{Full Sample} & \multicolumn{1}{c}{Resident}& \multicolumn{1}{c}{NSFH}& \multicolumn{1}{c}{NSFG} \\ 
\\[-1.8ex] & \multicolumn{1}{c}{(1)} & \multicolumn{1}{c}{(2)} & \multicolumn{1}{c}{(3)} & \multicolumn{1}{c}{(4)}\\ 
\hline \\[-1.8ex] 
\\[-2.2ex] & \multicolumn{4}{c}{Risk of Marriage relative to Cohabitation} \\  
 \hline \\[-1.8ex]
 Unilateral Divorce &  $-0.24^{***}$  &  $-0.25^{***}$  &  $-0.28^{***}$  &  $-0.28^{***}$  \\ 
  & ( 0.06 ) & ( 0.08 ) & ( 0.09 ) & ( 0.09 ) \\  
 \hline \\[-1.8ex]
 Average Relative Risk &  0.64  &  0.63  &  0.59  &  0.6  \\ 
 \hline \\[-1.8ex]
 \\[-2.2ex] & \multicolumn{4}{c}{Risk of Breakup relative to Cohabitation} \\  
 \hline \\[-1.8ex]
 Unilateral Divorce &  $-$0.19$^{***}$  &  $-$0.16$^{***}$  &  $-$0.08  &  $-$0.24$^{*}$  \\ 
  & ( 0.07 ) & ( 0.06 ) & ( 0.05 ) & ( 0.14 ) \\  
 \hline \\[-1.8ex]
 Average Relative Risk &  0.67  &  0.71  &  0.83  &  0.62  \\ 
 \hline \\[-1.8ex]
State Fixed effects & Yes & Yes & Yes & Yes \\ 
Year Fixed effects & Yes & Yes & Yes & Yes \\ 
Age Polynomial & Yes & Yes & Yes & Yes \\
Picewise Duration & Yes & Yes & Yes & Yes \\ 
\hline
Observations & \multicolumn{1}{c}{ 138012 } & \multicolumn{1}{c}{ 81920 } & \multicolumn{1}{c}{ 77826 } & \multicolumn{1}{c}{ 60186 } \\ 
\hline
Censored spells(\%) & \multicolumn{1}{c}{ 10.18 } & \multicolumn{1}{c}{ 10.98 } & \multicolumn{1}{c}{ 11.6 } & \multicolumn{1}{c}{ 8.38 } \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}