### 2020-10-01 welfare.tex

\begin{tabular}{cccc}
    \hline\midrule
    \multicolumn{2}{c}{\textbf{Female}}& \multicolumn{2}{c}{\textbf{Male}}\\
    \cmidrule(l){1-2}\cmidrule(l){3-4}
     Mutual Consent & Unilateral Divorce & Mutual Consent & Unilateral Divorce\\
     \cmidrule(l){1-4}
    \multicolumn{4}{c}{\textit{Life-Time utilities in $t=0$}}\\[3ex]
     -364.62 &-368.13 &-351.53 &-351.88 \\
    \cmidrule(l){1-4}
    \multicolumn{4}{c}{\textit{Welfare Losses with Unilateral Divorce}}\\[3ex]
    \multicolumn{2}{c}{\Chartgirls{1.0}}& \multicolumn{2}{c}{\Chartguys{0.2508214676889376}}\\[-0.15ex]
    \multicolumn{2}{c}{12933.66 \$}& \multicolumn{2}{c}{3244.04 \$}\\
     \hline\midrule\cmidrule(l){1-4} 
      \multicolumn{4}{c}{\textit{Life-Time utilities in $t=0$ when cohabitation is not in the choice set}}\\[3ex]
    -362.49 &-365.86 &-353.11 &-354.13 \\
    \cmidrule(l){1-4}
    \multicolumn{4}{c}{\textit{Welfare Losses with Unilateral Divorce}}\\[3ex]
    \multicolumn{2}{c}{\Chartgirls{1.0}}& \multicolumn{2}{c}{\Chartguys{0.7564234326824255}}\\[-0.15ex]
    \multicolumn{2}{c}{13783.62 \$}& \multicolumn{2}{c}{10426.25 \$}\\
    \hline\hline
    \end{tabular}