### 2020-10-27 wmar_mlogit.tex
prepare for overleaf

% Table created by stargazer v.5.2.2 by Marek Hlavac, Harvard University. E-mail: hlavac at fas.harvard.edu
% Date and time: dom, ott 25, 2020 - 09:33:20
\begingroup 
\footnotesize 
\begin{tabular}{@{\extracolsep{5pt}}lcccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{4}{c}{\textit{Dependent variable: Relative risk of cohabiting wrt to marrying}} \\
\cline{2-5} \\[-1.5ex]
 & Full Sample & Resident & Full Sample & Resident \\[-1ex]
\\[-1.4ex] & (1) & (2) & (3) & (4)\\ 
\hline \\[-1.8ex] 
 UnDiv    & 1.236$^{**}$ & 1.310$^{**}$ &   & \\ 
   & (0.106) & (0.141) &   &\\

 UnDiv*NoTit &   &   & 1.250$^{**}$ & 1.325$^{**}$ \\ 
  &   &  & (0.115) & (0.153) \\ 
  UnDiv*Tit &   &   & 1.130 & 1.313 \\ 
  &   &   & (0.179) & (0.273) \\ 
  Tit &   &   & 1.074 & 1.090 \\ 
  &  &  & (0.090) & (0.115) \\ 
 \hline \\[-1.8ex] 
State Fixed effects & Yes & Yes & Yes & Yes \\ 
Duration Polynomials & Yes & Yes & Yes & Yes \\ 
Demographic Controls & Yes & Yes & Yes & Yes \\ 
Observations & 112,697 & 70,882 & 112,697 & 70,882 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular} 
\endgroup