### 2020-11-05 graph_transitions2.tex

\begin{tikzpicture}[domain=0:1,scale=1]
\node [ellipse,fill=blue!20,minimum height=2.5cm,minimum width=7.8cm,draw, name=input,align=center] (sum) {\huge\bf Single};
\node [ellipse,fill=none,minimum height=0.5cm,minimum width=7.8cm, above right=1.5cm and 7.449cm of sum ] (controller) {\\ \huge \bf Breakup};
\node [ellipse,fill=blue!20,minimum height=2.5cm,minimum width=7.8cm,draw, below right=1.5cm and 7.449cm of controller,align=center] (system) {\huge \bf Cohabitation};
\draw [decoration={markings,mark=at position -0.95 with
    {\arrow[scale=5,>=stealth]{<}}},postaction={decorate}] (controller) --  node[above left,yshift=0.3cm]{\huge \bf}(system);
\node [output, right of=system] (output) {fe};
\node [ellipse,fill=blue!20,minimum height=2.5cm,minimum width=7.8cm,draw, below left=1.5cm and 7.449cm of system,align=center] (measurements) {\huge \bf {\color{blue!20}a}Marriage };
\node [ellipse,fill=none,minimum height=1.5cm,minimum width=4.2cm, left of=measurements, node distance=12.98cm,align=left] (divorce) {\huge \bf Divorce};
\draw [decoration={markings,mark=at position -0.92 with
    {\arrow[scale=5,>=stealth]{<}}},postaction={decorate}] (sum) -- node[above right,yshift=0.3cm]{\huge \bf } (controller);
\draw [decoration={markings,mark=at position 1 with
    {\arrow[scale=5,>=stealth]{>}}},postaction={decorate}] (sum) -- node[above]{\huge\bf} (system);
\draw [decoration={markings,mark=at position 0.99 with
    {\arrow[scale=5,>=stealth]{>}}},postaction={decorate}] (system) --node[above left]{\huge\bf } (measurements);
\draw [decoration={markings,mark=at position 0.99 with
    {\arrow[scale=5,>=stealth]{>}}},postaction={decorate}] (measurements) --node[above]{\huge\bf } (divorce);
\draw [decoration={markings,mark=at position 1 with
    {\arrow[scale=5,>=stealth]{>}}},postaction={decorate}] (sum) --node[above right]{\huge } (measurements);
\draw [decoration={markings,mark=at position 1 with
    {\arrow[scale=5,>=stealth]{>}}},postaction={decorate}] (divorce) --node[right]{\huge } (sum);
node [near end] {$$} (sum);
\end{tikzpicture}